---
title: "CUU Habitat Change - Analysis of statistics for flagging change"
date: "08/09/2020"
author: 'JNCC'
licence: 'MIT licence'
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
  library(sf)
library(ggplot2)
library(dplyr)
library(tidyr)
library(plotly)
library(DT)
library(shiny)
library(plotly)
```

## Example site - Insh Marshes

This report is to explore the statistics generated from theSentinel-2 data for the Insh Marshes and the different ways in which we can flag change.

```{r, include=F}
#load in the data and flag change
#plot differences
dirpath <- 'ChangeDetection/'
hab_stat <- read.csv(paste0(dirpath,"Statistics/InshMarshes_EUNIS/Seasonal_statistics/InshMarshes_NDVI_seasonal_changestats.txt"))
poly_stat <- read.csv(paste0(dirpath,"Statistics/InshMarshes_EUNIS/Seasonal_statistics/InshMarshes_NDVI_seasonal_stats.txt"))
polygons <- sf::st_read(paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_1_clean.shp")) %>% sf::st_drop_geometry()

#join the data
poly_hab <- poly_stat %>% dplyr::left_join(polygons[,c('Id','EUNIS_DESC')], by=c('ID'='Id')) 
hab_stat <- hab_stat %>% dplyr::select(-X) %>% rename(EUNIS_DESC = 'get.habclass.')
poly_all <- poly_hab %>% dplyr::left_join(hab_stat,by=c('seasonyear','EUNIS_DESC')) 

#order factor levels - again
##get unique years
all_lev <- stringr::str_split(levels(poly_all$seasonyear),"_",simplify=T) %>% data.frame()  %>% dplyr::mutate_if(is.character,as.factor)

#order by season and year
all_lev$X1 <- factor(all_lev$X1, levels = c('Spring','Summer','Autumn','Winter'),ordered=T)
lev_sort <- all_lev %>% dplyr::arrange(X1) %>% dplyr::arrange(X2) %>% tidyr::unite(yearseason, sep="_")
##sort levels in data
poly_all$seasonyear <- factor(poly_all$seasonyear, levels = lev_sort$yearseason)

```

## Seasonal statistics

### Polygons flagged as changed per SD range

The tables explore where change is flagged as being outside of a range of 1, 2 or 3 standard deviations from the mean for the habitat type. This is flagging both positive change above the sd range and negative change below the sd range.

```{r, echo=F}
##above 3 SD
polygon_change_1SD <- poly_all %>% dplyr::mutate(hab_mean_poschange = hab_mean+hab_meansd,
                                      hab_mean_negchange = hab_mean-hab_meansd,
                                      hab_median_poschange = hab_median+hab_mediansd,
                                      hab_median_negchange = hab_median-hab_mediansd,
                                      hab_min_poschange = hab_min+hab_minsd,
                                      hab_min_negchange = hab_min-hab_minsd,
                                      hab_max_poschange = hab_max+hab_maxsd,
                                      hab_max_negchange = hab_max-hab_maxsd,
                                      hab_Q1_poschange = hab_Q1+hab_Q1sd,
                                      hab_Q1_negchange = hab_Q1-hab_Q1sd,
                                      hab_Q3_poschange = hab_Q1+hab_Q3sd,
                                      hab_Q3_negchange = hab_Q1-hab_Q3sd) %>% 
   dplyr::mutate(meanchange = ifelse(mean <hab_mean_negchange |mean > hab_mean_poschange,1,0 ), 
         medianchange = ifelse(median <hab_median_negchange | median > hab_median_poschange,1,0 ),
         minchange = ifelse(min <hab_min_negchange | min>hab_min_poschange,1,0 ),
         maxchange = ifelse(max <hab_max_negchange | max >hab_max_poschange,1,0 ),
         Q1change = ifelse(Q1 <hab_Q1_negchange | Q1 >hab_Q1_poschange,1,0 ),
         Q3change = ifelse(Q3 <hab_Q3_negchange | Q3 >hab_Q3_poschange,1,0 ))
flagged_1SD <- polygon_change_1SD %>% dplyr::group_by(index, seasonyear, EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n(), meanchange=sum(meanchange),medianchange=sum(medianchange),Q1change=sum(Q1change),Q3change=sum(Q3change))

# flag those above or below 2SD as change

##above 2 SD
polygon_change_2SD <- poly_all %>% dplyr::mutate(hab_mean_poschange = hab_mean+(2*hab_meansd),
                                      hab_mean_negchange = hab_mean-(2*hab_meansd),
                                      hab_median_poschange = hab_median+(2*hab_mediansd),
                                      hab_median_negchange = hab_median-(2*hab_mediansd),
                                      hab_min_poschange = hab_min+(2*hab_minsd),
                                      hab_min_negchange = hab_min-(2*hab_minsd),
                                      hab_max_poschange = hab_max+(2*hab_maxsd),
                                      hab_max_negchange = hab_max-(2*hab_maxsd),
                                      hab_Q1_poschange = hab_Q1+(2*hab_Q1sd),
                                      hab_Q1_negchange = hab_Q1-(2*hab_Q1sd),
                                      hab_Q3_poschange = hab_Q1+(2*hab_Q3sd),
                                      hab_Q3_negchange = hab_Q1-(2*hab_Q3sd)) %>% 
  dplyr::mutate(meanchange = ifelse(mean <hab_mean_negchange |mean > hab_mean_poschange,1,0 ), 
         medianchange = ifelse(median <hab_median_negchange | median > hab_median_poschange,1,0 ),
         minchange = ifelse(min <hab_min_negchange | min>hab_min_poschange,1,0 ),
         maxchange = ifelse(max <hab_max_negchange | max >hab_max_poschange,1,0 ),
         Q1change = ifelse(Q1 <hab_Q1_negchange | Q1 >hab_Q1_poschange,1,0 ),
         Q3change = ifelse(Q3 <hab_Q3_negchange | Q3 >hab_Q3_poschange,1,0 ))
flagged_2SD <- polygon_change_2SD %>% dplyr::group_by(index, seasonyear, EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n(),meanchange=sum(meanchange),medianchange=sum(medianchange),Q1change=sum(Q1change),Q3change=sum(Q3change))

##above 3 SD
polygon_change_3SD <- poly_all %>% dplyr::mutate(hab_mean_poschange = hab_mean+(3*hab_meansd),
                                      hab_mean_negchange = hab_mean-(3*hab_meansd),
                                      hab_median_poschange = hab_median+(3*hab_mediansd),
                                      hab_median_negchange = hab_median-(3*hab_mediansd),
                                      hab_min_poschange = hab_min+(3*hab_minsd),
                                      hab_min_negchange = hab_min-(3*hab_minsd),
                                      hab_max_poschange = hab_max+(3*hab_maxsd),
                                      hab_max_negchange = hab_max-(3*hab_maxsd),
                                      hab_Q1_poschange = hab_Q1+(3*hab_Q1sd),
                                      hab_Q1_negchange = hab_Q1-(3*hab_Q1sd),
                                      hab_Q3_poschange = hab_Q1+(3*hab_Q3sd),
                                      hab_Q3_negchange = hab_Q1-(3*hab_Q3sd)) %>% 
  dplyr::mutate(meanchange = ifelse(mean <hab_mean_negchange |mean > hab_mean_poschange,1,0 ), 
         medianchange = ifelse(median <hab_median_negchange | median > hab_median_poschange,1,0 ),
         minchange = ifelse(min <hab_min_negchange | min>hab_min_poschange,1,0 ),
         maxchange = ifelse(max <hab_max_negchange | max >hab_max_poschange,1,0 ),
         Q1change = ifelse(Q1 <hab_Q1_negchange | Q1 >hab_Q1_poschange,1,0 ),
         Q3change = ifelse(Q3 <hab_Q3_negchange | Q3 >hab_Q3_poschange,1,0 ))
flagged_3SD <- polygon_change_3SD %>% dplyr::group_by(index, seasonyear, EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n(),meanchange=sum(meanchange),medianchange=sum(medianchange),Q1change=sum(Q1change),Q3change=sum(Q3change))

```


```{r,echo=F}
shiny::shinyApp(
  ui=shiny::fluidPage(
    shiny::selectInput('changeflag','Standard Deviations to flag change:',choices=c('1SD','2SD','3SD')),
    shiny::selectInput('habitat','Habitat type:',choices=c(unique(as.character(flagged_1SD$EUNIS_DESC)))),
    
    DT::dataTableOutput('habtable')
  ),
  server=function(input,output){
      
      output$habtable <- DT::renderDataTable({
            habitat_type <- input$habitat
            flagSD <- input$changeflag
            if(flagSD=='1SD'){
              hab_sum <- flagged_1SD %>% dplyr::filter(EUNIS_DESC==habitat_type)
              } else if(flagSD == '2SD'){
                hab_sum <- flagged_2SD %>% dplyr::filter(EUNIS_DESC==habitat_type)
              }else{
                hab_sum <- flagged_3SD %>% dplyr::filter(EUNIS_DESC==habitat_type)
              }
            hab_sum
    })
  },

  options = list(height = 700)
)

```



### Average habitat statistics per season per year 

This lets you view the statistics calculated per habitat type which each polygon is compared to. The first plot shows the 1SD and 2SD range, with the total number of polygons assessed that season and year shown as you hover over the mean.

The below plots demonstrate within each season and year, the distribution of polygon data as both a histogram per season year and seasonal boxplots across years.

```{r, echo=F}

hab_stat <- poly_all %>%  dplyr::select(index, seasonyear, EUNIS_DESC, hab_mean, hab_meansd,hab_median,hab_mediansd,hab_Q1,hab_Q1sd,hab_Q3,hab_Q3sd) %>% unique()

#get total polygons per habitat
 totalhab <- poly_all %>% dplyr::select(ID,seasonyear, EUNIS_DESC) %>% dplyr::group_by(seasonyear,EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n()) %>% dplyr::left_join(hab_stat,by=c('seasonyear','EUNIS_DESC'))

 
shiny::shinyApp(
  ui=shiny::fluidPage(
    shiny::selectInput('statistic','Statistic:',choices=c('mean','median','Q1','Q3')),
    shiny::selectInput('habitat','Habitat type:',choices=c(unique(as.character(totalhab$EUNIS_DESC)))),
    plotly::plotlyOutput('habgraph'),
    shiny::selectInput('seasonyearhist','Season Year:',choices=c(as.character(unique(totalhab$seasonyear)))),
     plotly::plotlyOutput('histplot'),
        shiny::selectInput('season','Season:',choices=c('Spring','Summer','Autumn','Winter')),    
    plotly::plotlyOutput('boxplot')
    
    ),
  server=function(input,output){
      
      output$habgraph <- plotly::renderPlotly({
            habitat_type <- input$habitat
            stat <- input$statistic
            hab_sum <- totalhab %>% dplyr::filter(EUNIS_DESC==habitat_type) %>% ungroup
      p <- ggplot2::ggplot(hab_sum)+
        ggplot2::geom_point(aes(x = seasonyear, y = get(paste0('hab_',stat)),group=totalpolygons), colour = "#3F9C35") +
        ggplot2::geom_line(aes(x = seasonyear, y = get(paste0('hab_',stat)),group=EUNIS_DESC), colour = "#3F9C35") +
        ggplot2::theme_classic()+
        ggplot2::labs(y = paste0(stat,"NDVI")) +
        ggplot2::theme(axis.title.x = element_blank(),axis.text.x=element_text(angle=90))
      p2 <- p + 
        ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))),ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
        ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)
      
      plotly::ggplotly(p2)
      
    })
      
      output$histplot <- plotly::renderPlotly({
        habitat_type <- input$habitat
        seasonyear_select <- input$seasonyearhist
        stat_select <- input$statistic
        poly_sum <- poly_all %>% dplyr::filter(seasonyear==seasonyear_select & EUNIS_DESC==habitat_type) %>% dplyr::select(ID,mean,sd,median,Q1,Q3,date)
        hist <-  ggplot2::ggplot(poly_sum, aes(x=get(stat_select))) +   ggplot2::geom_histogram(color="black", fill="white") +ggplot2::theme_classic() +
          ggplot2::labs(x = paste(stat_select,"NDVI"))
        plotly::ggplotly(hist)
  })
      
   output$boxplot <- plotly::renderPlotly({
         stat <- input$statistic
         season_select <- input$season
         habitat_type <- input$habitat
         seas_sum <- poly_all %>% dplyr::filter(EUNIS_DESC==habitat_type & stringr::str_detect(seasonyear,season_select))
         boxp <- ggplot2::ggplot(seas_sum,aes(x = seasonyear, y = get(stat),group=EUNIS_DESC)) +
           geom_boxplot() +
        ggplot2::theme_classic()+
        ggplot2::labs(y = paste0(stat,"NDVI")) +
        ggplot2::theme(axis.title.x = element_blank(),axis.text.x=element_text(angle=90))
        plotly::ggplotly(boxp)
      })   
   },
  options = list(height = 1600)
  )

```




### Exploring polygon changes

The plots below show seasonal polygon statistics compared to the mean seasonal habitat statistics, with ranges within 1SD (in grey) and 2SD (in red). Those listed below which are flagged as "change" are where a value for the season has fallen outside of the 2SD habitat range. 
Several polygons have been picked out which show change:531,14993,6907,14519,3344,3862

```{r, echo=F}

shiny::shinyApp(
  ui=shiny::fluidPage(
        shiny::numericInput('polyid','Polygon ID:',value=531,min=min(polygon_change_2SD$ID),max=max(polygon_change_2SD$ID)),
        shiny::selectInput('polystat','Statistic:',choices=c('mean','median','Q1','Q3')),
        plotly::plotlyOutput('polygraph'),
        DT::dataTableOutput('polytable')
  ),
  server=function(input,output){
    output$polygraph <- plotly::renderPlotly({
      stat <- input$polystat
      pID <- as.numeric(input$polyid)
      polygondat <- polygon_change_2SD %>% dplyr::filter(ID==pID) %>% dplyr::arrange(seasonyear)
      poly <- ggplot2::ggplot(polygondat)+
        ggplot2::geom_point(aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
        ggplot2::geom_line(aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
        ggplot2::geom_point(aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
        ggplot2::geom_line(aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
        ggplot2::theme_classic()+
        ggplot2::labs(y = paste0(stat,"NDVI")) +
        ggplot2::theme(axis.title.x = element_blank(),axis.text.x=element_text(angle=90))
      poly2 <- poly +
        geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))), ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
        geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)
    
    ggplotly(poly2) 
    })
    
    output$polytable <- DT::renderDataTable({
      stat <- input$polystat
      pID <- as.numeric(input$polyid)
      polygondat <- polygon_change_2SD %>% dplyr::filter(ID==pID) %>% dplyr::arrange(seasonyear)
      polygondat %>% 
        dplyr::filter(get(paste0(stat,'change'))==1) %>% dplyr::select(ID,index,seasonyear,date)
    })

  },
  options = list(height = 1000)
)

```



## Monthly Statistics

```{r, include=F}
#load in the data and flag change
#plot differences
dirpath <- 'ChangeDetection/'
hab_stat_month <- read.csv(paste0(dirpath,"Statistics/InshMarshes_EUNIS/Monthly_statistics/InshMarshes_NDVI_monthly_changestats.txt"))
poly_stat_month <- read.csv(paste0(dirpath,"Statistics/InshMarshes_EUNIS/Monthly_statistics/InshMarshes_NDVI_monthly_stats.txt"))
polygons <- sf::st_read(paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_1_clean.shp")) %>% sf::st_drop_geometry()

#join the data
poly_hab_month <- poly_stat_month %>% dplyr::left_join(polygons[,c('Id','EUNIS_DESC')], by=c('ID'='Id')) 
#create monthdate column
month_year <- poly_hab_month %>%
      dplyr::mutate(month=ifelse(month<10,as.character(paste0("0",month)),as.character(month))) %>%
      dplyr::mutate(year=as.character(year)) %>%
      dplyr::mutate(monthname = stringr::str_glue("{year}-{month}")) %>%
      dplyr::mutate(monthdate = as.factor(zoo::as.yearmon(monthname)))

hab_stat_month <- hab_stat_month %>% dplyr::select(-X) %>% rename(EUNIS_DESC = 'get.habclass.')
poly_all_monthly <- month_year %>% dplyr::left_join(hab_stat_month,by=c('monthdate','EUNIS_DESC')) 
poly_all_monthly$monthdate <- as.factor(poly_all_monthly$monthdate)

#order by month and year

monthTS <- tibble::tibble(expand.grid(month = 1:12, year = 2014:2020)) %>%
  dplyr::left_join(tibble::tibble(month = 1:12, monthname = month.abb)) %>%
  dplyr::group_by(year) %>%
  dplyr::mutate(FULLNAME = paste0(monthname, " ", year),
                plotOrder = dplyr::row_number()) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.factor, as.character)

monthTS$FULLNAME <- factor(monthTS$FULLNAME,ordered=T)
##sort levels in data
poly_all_monthly$monthdate <- factor(poly_all_monthly$monthdate, levels = monthTS$FULLNAME )

```

### Polygons flagged as changed per SD range

The tables explore where change is flagged as being outside of a range of 1, 2 or 3 standard deviations from the mean for the habitat type. This is flagging both positive change above the sd range and negative change below the sd range.

```{r, echo=F}
##above 3 SD
polygon_changemon_1SD <- poly_all_monthly %>% dplyr::mutate(hab_mean_poschange = hab_mean+hab_meansd,
                                      hab_mean_negchange = hab_mean-hab_meansd,
                                      hab_median_poschange = hab_median+hab_mediansd,
                                      hab_median_negchange = hab_median-hab_mediansd,
                                      hab_min_poschange = hab_min+hab_minsd,
                                      hab_min_negchange = hab_min-hab_minsd,
                                      hab_max_poschange = hab_max+hab_maxsd,
                                      hab_max_negchange = hab_max-hab_maxsd,
                                      hab_Q1_poschange = hab_Q1+hab_Q1sd,
                                      hab_Q1_negchange = hab_Q1-hab_Q1sd,
                                      hab_Q3_poschange = hab_Q1+hab_Q3sd,
                                      hab_Q3_negchange = hab_Q1-hab_Q3sd) %>% 
   dplyr::mutate(meanchange = ifelse(mean <hab_mean_negchange |mean > hab_mean_poschange,1,0 ), 
         medianchange = ifelse(median <hab_median_negchange | median > hab_median_poschange,1,0 ),
         minchange = ifelse(min <hab_min_negchange | min>hab_min_poschange,1,0 ),
         maxchange = ifelse(max <hab_max_negchange | max >hab_max_poschange,1,0 ),
         Q1change = ifelse(Q1 <hab_Q1_negchange | Q1 >hab_Q1_poschange,1,0 ),
         Q3change = ifelse(Q3 <hab_Q3_negchange | Q3 >hab_Q3_poschange,1,0 ))
flaggedmon_1SD <- polygon_changemon_1SD %>% dplyr::group_by(index, monthdate, EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n(), meanchange=sum(meanchange),medianchange=sum(medianchange),Q1change=sum(Q1change),Q3change=sum(Q3change))

# flag those above or below 2SD as change

##above 2 SD
polygon_changemon_2SD <- poly_all_monthly %>% dplyr::mutate(hab_mean_poschange = hab_mean+(2*hab_meansd),
                                      hab_mean_negchange = hab_mean-(2*hab_meansd),
                                      hab_median_poschange = hab_median+(2*hab_mediansd),
                                      hab_median_negchange = hab_median-(2*hab_mediansd),
                                      hab_min_poschange = hab_min+(2*hab_minsd),
                                      hab_min_negchange = hab_min-(2*hab_minsd),
                                      hab_max_poschange = hab_max+(2*hab_maxsd),
                                      hab_max_negchange = hab_max-(2*hab_maxsd),
                                      hab_Q1_poschange = hab_Q1+(2*hab_Q1sd),
                                      hab_Q1_negchange = hab_Q1-(2*hab_Q1sd),
                                      hab_Q3_poschange = hab_Q1+(2*hab_Q3sd),
                                      hab_Q3_negchange = hab_Q1-(2*hab_Q3sd)) %>% 
  dplyr::mutate(meanchange = ifelse(mean <hab_mean_negchange |mean > hab_mean_poschange,1,0 ), 
         medianchange = ifelse(median <hab_median_negchange | median > hab_median_poschange,1,0 ),
         minchange = ifelse(min <hab_min_negchange | min>hab_min_poschange,1,0 ),
         maxchange = ifelse(max <hab_max_negchange | max >hab_max_poschange,1,0 ),
         Q1change = ifelse(Q1 <hab_Q1_negchange | Q1 >hab_Q1_poschange,1,0 ),
         Q3change = ifelse(Q3 <hab_Q3_negchange | Q3 >hab_Q3_poschange,1,0 ))
flaggedmon_2SD <- polygon_changemon_2SD %>% dplyr::group_by(index, monthdate, EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n(),meanchange=sum(meanchange),medianchange=sum(medianchange),Q1change=sum(Q1change),Q3change=sum(Q3change))

##above 3 SD
polygon_changemon_3SD <- poly_all_monthly %>% dplyr::mutate(hab_mean_poschange = hab_mean+(3*hab_meansd),
                                      hab_mean_negchange = hab_mean-(3*hab_meansd),
                                      hab_median_poschange = hab_median+(3*hab_mediansd),
                                      hab_median_negchange = hab_median-(3*hab_mediansd),
                                      hab_min_poschange = hab_min+(3*hab_minsd),
                                      hab_min_negchange = hab_min-(3*hab_minsd),
                                      hab_max_poschange = hab_max+(3*hab_maxsd),
                                      hab_max_negchange = hab_max-(3*hab_maxsd),
                                      hab_Q1_poschange = hab_Q1+(3*hab_Q1sd),
                                      hab_Q1_negchange = hab_Q1-(3*hab_Q1sd),
                                      hab_Q3_poschange = hab_Q1+(3*hab_Q3sd),
                                      hab_Q3_negchange = hab_Q1-(3*hab_Q3sd)) %>% 
  dplyr::mutate(meanchange = ifelse(mean <hab_mean_negchange |mean > hab_mean_poschange,1,0 ), 
         medianchange = ifelse(median <hab_median_negchange | median > hab_median_poschange,1,0 ),
         minchange = ifelse(min <hab_min_negchange | min>hab_min_poschange,1,0 ),
         maxchange = ifelse(max <hab_max_negchange | max >hab_max_poschange,1,0 ),
         Q1change = ifelse(Q1 <hab_Q1_negchange | Q1 >hab_Q1_poschange,1,0 ),
         Q3change = ifelse(Q3 <hab_Q3_negchange | Q3 >hab_Q3_poschange,1,0 ))
flaggedmon_3SD <- polygon_changemon_3SD %>% dplyr::group_by(index, monthdate, EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n(),meanchange=sum(meanchange),medianchange=sum(medianchange),Q1change=sum(Q1change),Q3change=sum(Q3change))


```


```{r,echo=F}
shiny::shinyApp(
  ui=shiny::fluidPage(
    shiny::selectInput('changeflag','Standard Deviations to flag change:',choices=c('1SD','2SD','3SD')),
    shiny::selectInput('habitat','Habitat type:',choices=c(unique(as.character(flagged_1SD$EUNIS_DESC)))),
    
    DT::dataTableOutput('habtable')
  ),
  server=function(input,output){
      
      output$habtable <- DT::renderDataTable({
            habitat_type <- input$habitat
            flagSD <- input$changeflag
            if(flagSD=='1SD'){
              hab_sum <- flaggedmon_1SD %>% dplyr::filter(EUNIS_DESC==habitat_type)
              } else if(flagSD == '2SD'){
                hab_sum <- flaggedmon_2SD %>% dplyr::filter(EUNIS_DESC==habitat_type)
              }else{
                hab_sum <- flaggedmon_3SD %>% dplyr::filter(EUNIS_DESC==habitat_type)
              }
            hab_sum
    })
  },

  options = list(height = 700)
)

```



### Average habitat statistics per month per year 

This lets you view the statistics calculated per habitat type which each polygon is compared to. The first plot shows the 1SD and 2SD range, with the total number of polygons assessed that month and year shown as you hover over the mean.

The below plots demonstrate within each month and year, the distribution of polygon data as both a histogram per month year and monthly boxplots across years.

```{r, echo=F}

hab_statmon <- poly_all_monthly %>%  dplyr::select(index, monthdate, EUNIS_DESC, hab_mean, hab_meansd,hab_median,hab_mediansd,hab_Q1,hab_Q1sd,hab_Q3,hab_Q3sd) %>% unique()

#get total polygons per habitat
 totalhabmon <- poly_all_monthly %>% dplyr::select(ID,monthdate, EUNIS_DESC) %>% dplyr::group_by(monthdate,EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n()) %>% dplyr::left_join(hab_statmon,by=c('monthdate','EUNIS_DESC')) 
 
shiny::shinyApp(
  ui=shiny::fluidPage(
    shiny::selectInput('statistic','Statistic:',choices=c('mean','median','Q1','Q3')),
    shiny::selectInput('habitat','Habitat type:',choices=c(unique(as.character(totalhabmon$EUNIS_DESC)))),
    plotly::plotlyOutput('habgraph'),
    shiny::selectInput('monthyearhist','Month Year:',choices=c(as.character(unique(totalhabmon$monthdate)))),
     plotly::plotlyOutput('histplot'),
        shiny::selectInput('month','Month:',choices=c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec')),    
    plotly::plotlyOutput('boxplot')
    

  ),
  server=function(input,output){
      
      output$habgraph <- plotly::renderPlotly({
            habitat_type <- input$habitat
            stat <- input$statistic
            hab_sum <- totalhabmon %>% dplyr::filter(EUNIS_DESC==habitat_type)
      p <- ggplot2::ggplot(hab_sum)+
        ggplot2::geom_point(aes(x = monthdate, y = get(paste0('hab_',stat)),group=totalpolygons), colour = "#3F9C35") +
        ggplot2::geom_line(aes(x = monthdate, y = get(paste0('hab_',stat)),group=EUNIS_DESC), colour = "#3F9C35") +
        ggplot2::theme_classic()+
        ggplot2::labs(y = paste0(stat,"NDVI")) +
        ggplot2::theme(axis.title.x = element_blank(),axis.text.x=element_text(angle=90))
      p2 <- p + 
        ggplot2::geom_ribbon(aes(x = 1:length(monthdate), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))),ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
        ggplot2::geom_ribbon(aes(x = 1:length(monthdate), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)
      
      plotly::ggplotly(p2)
      
    })
      
      output$histplot <- plotly::renderPlotly({
        habitat_type <- input$habitat
        monthyear_select <- input$monthyearhist
        stat_select <- input$statistic
        poly_sum <- poly_all_monthly %>% dplyr::filter(monthdate==monthyear_select & EUNIS_DESC==habitat_type) %>% dplyr::select(ID,mean,sd,median,Q1,Q3,date)
        hist <-  ggplot2::ggplot(poly_sum, aes(x=get(stat_select))) +   ggplot2::geom_histogram(color="black", fill="white") +ggplot2::theme_classic() +
          ggplot2::labs(x = paste(stat_select,"NDVI"))
        plotly::ggplotly(hist)
  })
      
   output$boxplot <- plotly::renderPlotly({
         stat <- input$statistic
         month_select <- input$month
         habitat_type <- input$habitat
         seas_sum <- poly_all_monthly %>% dplyr::filter(EUNIS_DESC==habitat_type & stringr::str_detect(monthdate,month_select))
         boxp <- ggplot2::ggplot(seas_sum,aes(x = monthdate, y = get(stat),group=EUNIS_DESC)) +
           geom_boxplot() +
        ggplot2::theme_classic()+
        ggplot2::labs(y = paste0(stat,"NDVI")) +
        ggplot2::theme(axis.title.x = element_blank(),axis.text.x=element_text(angle=90))
        plotly::ggplotly(boxp)
      })   
      
      
 },
  options = list(height = 1600)
)



```



### Exploring polygon changes

The plots below show monthly polygon statistics compared to the mean monthly habitat statistics, with ranges within 1SD (in grey) and 2SD (in red). Those listed below which are flagged as "change" are where a value for the month has fallen outside of the 2SD habitat range. 

Several polygons have been picked out which show change:531,14993,6907,14519,3344,3862

```{r, echo=F}

shiny::shinyApp(
  ui=shiny::fluidPage(
        shiny::numericInput('polyid','Polygon ID:',value=531,min=min(polygon_changemon_2SD$ID),max=max(polygon_changemon_2SD$ID)),
        shiny::selectInput('polystat','Statistic:',choices=c('mean','median','Q1','Q3')),
        plotly::plotlyOutput('polygraph'),
        DT::dataTableOutput('polytable')
  ),
  server=function(input,output){
    output$polygraph <- plotly::renderPlotly({
      stat <- input$polystat
      pID <- as.numeric(input$polyid)
      polygondat <- polygon_changemon_2SD %>% dplyr::filter(ID==pID) %>% dplyr::arrange(monthdate)
      poly <- ggplot2::ggplot(polygondat)+
        ggplot2::geom_point(aes(x = monthdate, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
        ggplot2::geom_line(aes(x = monthdate, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
        ggplot2::geom_point(aes(x = monthdate, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
        ggplot2::geom_line(aes(x = monthdate, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
        ggplot2::theme_classic()+
        ggplot2::labs(y = paste0(stat,"NDVI")) +
        ggplot2::theme(axis.title.x = element_blank(),axis.text.x=element_text(angle=90))
      poly2 <- poly +
        geom_ribbon(aes(x = 1:length(monthdate), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))), ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
        geom_ribbon(aes(x = 1:length(monthdate), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)
    
    ggplotly(poly2) 
    })
    
    output$polytable <- DT::renderDataTable({
            pID <- as.numeric(input$polyid)
      stat <- input$polystat
            polygondat <- polygon_changemon_2SD %>% dplyr::filter(ID==pID) %>% dplyr::arrange(monthdate)
      polygondat %>% 
        dplyr::filter(get(paste0(stat,'change'))==1) %>% dplyr::select(ID,index,monthdate,date)
    })

  },
  options = list(height = 1300)
)

```


