---
title: "CUU Habitat Change - Analysis of change flag"
author: "Becky Trippier"
date: "12/08/2020"
output:
  html_document:
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
  library(sf)
library(ggplot2)
library(dplyr)
library(tidyr)
library(plotly)
library(DT)
```

## Example site - Insh Marshes

```{r, include=F,eval=F}
#load in the data and flag change
#plot differences
dirpath <- 'Scotland/'
hab_stat <- read.csv(paste0(dirpath,"Statistics/InshMarshes_EUNIS/Seasonal_statistics/InshMarshes_NDVI_seasonal_changestats.txt"))
poly_stat <- read.csv(paste0(dirpath,"Statistics/InshMarshes_EUNIS/Seasonal_statistics/InshMarshes_NDVI_seasonal_stats.txt"))
polygons <- sf::st_read(paste0(dirpath,"Data/Habitat_maps/Scotland/ELCS_Scottish_site_1.shp")) %>% sf::st_drop_geometry()

#join the data
poly_hab <- poly_stat %>% dplyr::left_join(polygons[,c('Id','EUNIS_DESC')], by=c('ID'='Id')) 
hab_stat <- hab_stat %>% dplyr::select(-X) %>% dplyr::rename(EUNIS_DESC = 'get.habclass.')
poly_all <- poly_hab %>% dplyr::left_join(hab_stat,by=c('seasonyear','EUNIS_DESC')) 

#order factor levels - again
##get unique years
all_lev <- stringr::str_split(levels(poly_all$seasonyear),"_",simplify=T) %>% data.frame()  %>% dplyr::mutate_if(is.character,as.factor)

#order by season and year
all_lev$X1 <- factor(all_lev$X1, levels = c('Spring','Summer','Autumn','Winter'),ordered=T)
lev_sort <- all_lev %>% arrange(X1) %>% dplyr::arrange(X2) %>% tidyr::unite(yearseason, sep="_")
##sort levels in data
poly_all$seasonyear <- factor(poly_all$seasonyear, levels = lev_sort$yearseason)

# flag those above or below 2SD as change

##above 2 SD
polygon_change <- poly_all %>% dplyr::mutate(hab_mean_poschange = hab_mean+(2*hab_meansd),
                                      hab_mean_negchange = hab_mean-(2*hab_meansd),
                                      hab_median_poschange = hab_median+(2*hab_mediansd),
                                      hab_median_negchange = hab_median-(2*hab_mediansd),
                                      hab_min_poschange = hab_min+(2*hab_minsd),
                                      hab_min_negchange = hab_min-(2*hab_minsd),
                                      hab_max_poschange = hab_max+(2*hab_maxsd),
                                      hab_max_negchange = hab_max-(2*hab_maxsd),
                                      hab_Q1_poschange = hab_Q1+(2*hab_Q1sd),
                                      hab_Q1_negchange = hab_Q1-(2*hab_Q1sd),
                                      hab_Q3_poschange = hab_Q1+(2*hab_Q3sd),
                                      hab_Q3_negchange = hab_Q1-(2*hab_Q3sd)) %>% 
  dplyr::mutate(meanchange = ifelse(mean <hab_mean_negchange,1,0 ), 
         medianchange = ifelse(median <hab_median_negchange,1,0 ),
         minchange = ifelse(min <hab_min_negchange,1,0 ),
         maxchange = ifelse(max <hab_max_negchange,1,0 ),
         Q1change = ifelse(Q1 <hab_Q1_negchange,1,0 ),
         Q3change = ifelse(Q3 <hab_Q3_negchange,1,0 ))



```

## Numbers of polygons flagged as change

### When change is calculated by 1 standard deviations from the mean for the habitat type:

```{r, echo=F,eval=F}
##above 3 SD
polygon_change_1SD <- poly_all %>% dplyr:: mutate(hab_mean_poschange = hab_mean+hab_meansd,
                                      hab_mean_negchange = hab_mean-hab_meansd,
                                      hab_median_poschange = hab_median+hab_mediansd,
                                      hab_median_negchange = hab_median-hab_mediansd,
                                      hab_min_poschange = hab_min+hab_minsd,
                                      hab_min_negchange = hab_min-hab_minsd,
                                      hab_max_poschange = hab_max+hab_maxsd,
                                      hab_max_negchange = hab_max-hab_maxsd,
                                      hab_Q1_poschange = hab_Q1+hab_Q1sd,
                                      hab_Q1_negchange = hab_Q1-hab_Q1sd,
                                      hab_Q3_poschange = hab_Q1+hab_Q3sd,
                                      hab_Q3_negchange = hab_Q1-hab_Q3sd) %>% 
   dplyr::mutate(meanchange = ifelse(mean <hab_mean_negchange,1,0 ), 
         medianchange = ifelse(median <hab_median_negchange,1,0 ),
         minchange = ifelse(min <hab_min_negchange,1,0 ),
         maxchange = ifelse(max <hab_max_negchange,1,0 ),
         Q1change = ifelse(Q1 <hab_Q1_negchange,1,0 ),
         Q3change = ifelse(Q3 <hab_Q3_negchange,1,0 ))
#flagged summary
flagged_1SD <- polygon_change_1SD %>% dplyr::group_by(index, seasonyear, EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n(), meanchange=sum(meanchange),medianchange=sum(medianchange),Q1change=sum(Q1change),Q3change=sum(Q3change))
datatable(flagged_1SD)
```

### When change is calculated by 2 standard deviations from the mean for the habitat type:

```{r, echo=F,eval=F}
flagged <- polygon_change %>% dplyr::group_by(index, seasonyear, EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n(),meanchange=sum(meanchange),medianchange=sum(medianchange),Q1change=sum(Q1change),Q3change=sum(Q3change))
datatable(flagged)
```

### When change is calculated by 3 standard deviations from the mean for the habitat type:

```{r, echo=F,eval=F}
##above 3 SD
polygon_change_3SD <- poly_all %>% dplyr::mutate(hab_mean_poschange = hab_mean+(3*hab_meansd),
                                      hab_mean_negchange = hab_mean-(3*hab_meansd),
                                      hab_median_poschange = hab_median+(3*hab_mediansd),
                                      hab_median_negchange = hab_median-(3*hab_mediansd),
                                      hab_min_poschange = hab_min+(3*hab_minsd),
                                      hab_min_negchange = hab_min-(3*hab_minsd),
                                      hab_max_poschange = hab_max+(3*hab_maxsd),
                                      hab_max_negchange = hab_max-(3*hab_maxsd),
                                      hab_Q1_poschange = hab_Q1+(3*hab_Q1sd),
                                      hab_Q1_negchange = hab_Q1-(3*hab_Q1sd),
                                      hab_Q3_poschange = hab_Q1+(3*hab_Q3sd),
                                      hab_Q3_negchange = hab_Q1-(3*hab_Q3sd)) %>% 
  dplyr::mutate(meanchange = ifelse(mean <hab_mean_negchange,1,0 ), 
         medianchange = ifelse(median <hab_median_negchange,1,0 ),
         minchange = ifelse(min <hab_min_negchange,1,0 ),
         maxchange = ifelse(max <hab_max_negchange,1,0 ),
         Q1change = ifelse(Q1 <hab_Q1_negchange,1,0 ),
         Q3change = ifelse(Q3 <hab_Q3_negchange,1,0 ))
#flagged summary
flagged_3SD <- polygon_change_3SD %>% dplyr::group_by(index, seasonyear, EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n(),meanchange=sum(meanchange),medianchange=sum(medianchange),Q1change=sum(Q1change),Q3change=sum(Q3change))
datatable(flagged_3SD)
```


## Average habitat statistics per month per season


```{r, echo=F,eval=F}
hab_stat <- polygon_change %>%  
  dplyr::select(index, seasonyear, EUNIS_DESC, hab_mean, hab_meansd,hab_median,hab_mediansd,hab_Q1,hab_Q1sd,hab_Q3,hab_Q3sd) %>% 
  unique()

#get total polygons per habitat
 totalhab <- polygon_change %>% 
   dplyr::select(ID,seasonyear, EUNIS_DESC) %>% 
   dplyr::group_by(seasonyear,EUNIS_DESC) %>% dplyr::summarise(totalpolygons=n()) %>% 
   dplyr::left_join(hab_stat,by=c('seasonyear','EUNIS_DESC'))


datatable(totalhab)

```




## Exploring polygon changes

The plots below show seasonal polygon statistics compared to the mean seasonal habitat statistics, with ranges within 1SD (in grey) and 2SD (in red). Those listed below which are flagged as "change" are where a value for the season has fallen below the lower habitat range.

### Polygon id: 531

```{r, echo=F}
#---------------------#
#filter to selected polygon and stat
poly_id = 531
stat <-'mean'
#---------------------#

polygondat <- polygon_change %>% dplyr::filter(ID==poly_id)  %>% dplyr::arrange(seasonyear)

p <- ggplot2::ggplot(polygondat)+
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::theme_classic()+
  ggplot2::labs(y = paste0(stat,"NDVI")) +
  ggplot2::theme(axis.title.x = ggplot2::element_blank(),axis.text.x=ggplot2::element_text(angle=90))

#separate as geom_ribbon takes continuous data
p2 <- p + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))), ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)

plotly::ggplotly(p2) 

```


```{r, echo=F}
# polygons flagged as change
polygondat %>% dplyr::filter(get(paste0(stat,'change'))==1) %>% dplyr::select(ID,index,seasonyear,date)

```

### Polygon id: 14993

```{r, echo=F}
#---------------------#
#filter to selected polygon and stat
poly_id = 14993
stat <-'mean'
#---------------------#

polygondat <- polygon_change %>% dplyr::filter(ID==poly_id)  %>% dplyr::arrange(seasonyear)

p <- ggplot2::ggplot(polygondat)+
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::theme_classic()+
  ggplot2::labs(y = paste0(stat,"NDVI")) +
  ggplot2::theme(axis.title.x = ggplot2::element_blank(),axis.text.x=ggplot2::element_text(angle=90))

#separate as geom_ribbon takes continuous data
p2 <- p + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))), ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)

plotly::ggplotly(p2) 

```


```{r, echo=F}
# polygons flagged as change
polygondat %>% dplyr::filter(get(paste0(stat,'change'))==1) %>% dplyr::select(ID,index,seasonyear,date)
```


### Polygon id: 6907

```{r, echo=F}
#---------------------#
#filter to selected polygon and stat
poly_id = 6907
stat <-'mean'
#---------------------#

polygondat <- polygon_change %>% dplyr::filter(ID==poly_id)  %>% dplyr::arrange(seasonyear)

p <- ggplot2::ggplot(polygondat)+
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::theme_classic()+
  ggplot2::labs(y = paste0(stat,"NDVI")) +
  ggplot2::theme(axis.title.x = ggplot2::element_blank(),axis.text.x=ggplot2::element_text(angle=90))

#separate as geom_ribbon takes continuous data
p2 <- p + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))), ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)

plotly::ggplotly(p2) 

```


```{r, echo=F}
# polygons flagged as change
polygondat %>% dplyr::filter(get(paste0(stat,'change'))==1) %>% dplyr::select(ID,index,seasonyear,date)
```

### Polygon id: 14519

```{r, echo=F}
#---------------------#
#filter to selected polygon and stat
poly_id = 14519
stat <-'mean'
#---------------------#

polygondat <- polygon_change %>% dplyr::filter(ID==poly_id)  %>% dplyr::arrange(seasonyear)

p <- ggplot2::ggplot(polygondat)+
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::theme_classic()+
  ggplot2::labs(y = paste0(stat,"NDVI")) +
  ggplot2::theme(axis.title.x = ggplot2::element_blank(),axis.text.x=ggplot2::element_text(angle=90))

#separate as geom_ribbon takes continuous data
p2 <- p + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))), ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)

plotly::ggplotly(p2) 

```


```{r, echo=F}
# polygons flagged as change
polygondat %>% dplyr::filter(get(paste0(stat,'change'))==1) %>% dplyr::select(ID,index,seasonyear,date)
```

### Polygon id: 3344

```{r,echo=F}
#---------------------#
#filter to selected polygon and stat
poly_id = 3344
stat <-'mean'
#---------------------#

polygondat <- polygon_change %>% dplyr::filter(ID==poly_id)  %>% dplyr::arrange(seasonyear)

p <- ggplot2::ggplot(polygondat)+
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::theme_classic()+
  ggplot2::labs(y = paste0(stat,"NDVI")) +
  ggplot2::theme(axis.title.x = ggplot2::element_blank(),axis.text.x=ggplot2::element_text(angle=90))

#separate as geom_ribbon takes continuous data
p2 <- p + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))), ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)

plotly::ggplotly(p2) 

```


```{r, echo=F}
# polygons flagged as change
polygondat %>% dplyr::filter(get(paste0(stat,'change'))==1) %>% dplyr::select(ID,index,seasonyear,date)
```

### Polygon id: 3862

```{r, echo=F}
#---------------------#
#filter to selected polygon and stat
poly_id = 3862
stat <-'mean'
#---------------------#

polygondat <- polygon_change %>% dplyr::filter(ID==poly_id)  %>% dplyr::arrange(seasonyear)

p <- ggplot2::ggplot(polygondat)+
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(stat), group=EUNIS_DESC), colour = "#1E1E1E") +
  ggplot2::geom_point(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::geom_line(ggplot2::aes(x = seasonyear, y = get(paste0('hab_',stat)), group=EUNIS_DESC), colour = "#3F9C35") +
  ggplot2::theme_classic()+
  ggplot2::labs(y = paste0(stat,"NDVI")) +
  ggplot2::theme(axis.title.x = ggplot2::element_blank(),axis.text.x=ggplot2::element_text(angle=90))

#separate as geom_ribbon takes continuous data
p2 <- p + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-(2*get(paste0('hab_',stat,'sd'))), ymax=get(paste0('hab_',stat))+(2*get(paste0('hab_',stat,'sd')))), fill="red",alpha=0.1) + 
  ggplot2::geom_ribbon(aes(x = 1:length(seasonyear), ymin = get(paste0('hab_',stat))-get(paste0('hab_',stat,'sd')), ymax=get(paste0('hab_',stat))+get(paste0('hab_',stat,'sd'))), fill="grey70",alpha=0.5)

plotly::ggplotly(p2) 

```


```{r, echo=F}
# polygons flagged as change
polygondat %>% dplyr::filter(get(paste0(stat,'change'))==1) %>% dplyr::select(ID,index,seasonyear,date)
```
