---
title: "CUU_ChangeDetection Welsh sites"
date: "23/07/2020"
output:
  html_document:
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library(HabChangeDetection)
library(sf)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(tmap)
library(janitor)
library(fs)
library(readxl)
library(magrittr)

source('habitat_change_detection_functions.R')

```

```{r display_site,include=F,eval=F}
display_site <- function(site_polygons, bounding_box) {
  tmap::tmap_mode("view")
  tmap::tm_shape(bounding_box) +
    tmap::tm_fill(col = "grey", alpha = 0.5) +
    tmap::tm_borders(col = "red", lwd = 2) +
    tmap::tm_shape(site_polygons) +
    tmap::tm_fill(col = "grey", alpha = 0.8) +
    tmap::tm_borders(col = "black", lwd = 1) +
    tmap::tm_basemap(leaflet::providers$OpenStreetMap) +
    tmap::tm_scale_bar()
}
```

```{r display_table, echo=FALSE,include=F}
library(reactable)
display_table <- function(data, rows = nrow(data)) {
  reactable::reactable(data[1:rows, ], fullWidth = FALSE, searchable = TRUE, 
            compact = TRUE, wrap = FALSE, resizable = TRUE,
            defaultColDef = colDef(align = "left"),
            showPageSizeOptions = TRUE, pageSizeOptions = c(10, nrow(data)),
            class = "table")
}
```

```{r get_granule_present_month,include=F}
# at least one S2 granules present per month
get_granule_present_month <- function(granules_dataset){
  granules_dataset %>% 
    dplyr::mutate(year = year(capture_date),
                  month = month(capture_date)) %>% 
    dplyr::group_by(year, month) %>% 
    dplyr::tally() %>% 
    dplyr::full_join(tidyr::crossing(year = seq(2015, 2020), month = seq(1, 12))) %>% 
    dplyr::mutate(n = tidyr::replace_na(n, 0)) %>% 
    dplyr::arrange(year, month) %>% 
    dplyr::mutate(month = month.abb[month]) %>% 
    tidyr::pivot_wider(names_from = month, values_from = n) %>% 
    dplyr::mutate(across(.cols = Jan:Dec, .fns = function(x){if(x > 0){ x / x } else 0})) %>% 
    janitor::adorn_totals(where = "col")
}
```

```{r paths, include=F,eval=F}
site_folder_path <- path("Sites/")
s2_folder_path <- path("ChangeDetection/Data/Sentinel2/")
s1_folder_path <- path("ChangeDetection/Data/Sentinel1/")
hab_map_path<-path("ChangeDetection/Data/Habitat_maps/Wales/Phase_1_corrected_20200814/")
```

## Copernicus User Uptake - Change detection

Welsh sites:

* Cors Bodeilio site (10km^2^)  
* Rhos Tonyrefail square site (10km^2^)

Spatial frameworks:

* Phase 1 Habitat Map of Wales
* Finer scale UKHab classification segmentation


### Cors Bodeilio site (10km^2^) 

**S2 Granule:** T30UVE

* Cors Bodeilio (fen)

```{r corssite,include=FALSE, fig.width=8, fig.height=4,eval=F}
cors_bodeilio_site <- sf::st_read(path(site_folder_path, "CorsBodeilio.shp")) %>% 
  janitor::clean_names() %>% 
  dplyr::mutate(name = "Cors Bodeilio",
         code = 1)
cors_bodeilio_bounding_box <- create_square_bounding_box_polygon(cors_bodeilio_site)
# sf::st_write(cors_bodeilio_bounding_box, 
#        path(site_folder_path, "bounding_box", "cors_bodeilio_bounding_box.kml"))
display_site(cors_bodeilio_site, cors_bodeilio_bounding_box)

```

### S1 and S2 Granules

### Cleaning the Spatial framework

The Cors Bodeilio Phase 1 shapefile was cleaned by removing polygons of <0.01 ha (100m2) and unwanted habitat covers. A total of 87 polygons were removed.   

```{r clean_corsbodeilio_Phase1_habitat_layer,eval=FALSE}
cors_raw<-sf::st_read("CorsBodeilio_Phase1_names.shp")
cors_clean<-cors_raw %>% dplyr::filter(!Name %in% c("Intertidal mud/sand","Intertidal cobbles/shingle","Intertidal rocks/boulders","Salt marsh","Mud/sand above mhw","Shingle/gravel above mhw","Rocks/boulders above mhw","Open dune","Coastal grassland")) %>% dplyr::filter(AREA_HA>=0.01)

st_write(cors_clean,"Cleaned/CorsBodeilio_Phase1_names_cleaned.shp")
```

### Running the zonal stats
```{r run_zonal_stats_corsbodeilio,eval=FALSE}
dirpath<-"ChangeDetection/"

zonal_stats(polys=paste0(dirpath,'Data/Habitat_maps/Wales/Phase_1_corrected_20200814/Cleaned/CorsBodeilio_Phase1_names_cleaned.shp'),
            polyfield='OBJECTID',
            s2path=paste0(dirpath,'Data/Sentinel2/CorsBodeilio/'),
            s1path=paste0(dirpath,'Data/Sentinel1/CorsBodeilio/'),
            outfolder=paste0(dirpath,'Statistics/CorsBodeilio_Phase1/'),
            sitename='CorsBodeilio')
            
```

### Calculate monthly and seasonal summary stats and change stats

```{r run_summary_change_stats_corsbodeilio,eval=FALSE}
change_stats(sitefile = paste0(dirpath,'Statistics/CorsBodeilio_Phase1/ZonalStats/CorsBodeilio_zonal_stats.txt'),
             outfolder = paste0(dirpath,'Statistics/CorsBodeilio_Phase1/'),
             polygons = paste0(dirpath,'Data/Habitat_maps/Wales/Phase_1_corrected_20200814/Cleaned/CorsBodeilio_Phase1_names_cleaned.shp'),
             polyid = 'OBJECTID',
             habclass = 'Name')
```


### Rhos Tonyrefail square site (10km^2^): 

**S2 Granule:** T30UVC

* Rhos Tonyrefail square (marshy and neutral grassland, possible small amount of peat)

```{r Rhossite, include=FALSE, fig.width=8, fig.height=4}
rhos_tonyrefail_site <- sf::st_read(path(site_folder_path, "Rhos_Tonyrefail_square.shp")) %>% 
  janitor::clean_names() %>% 
  dplyr::rename(name = site_name,
         code = tile_name) %>% 
  dplyr::select(name, code)
rhos_tonyrefail_bounding_box <- create_square_bounding_box_polygon(rhos_tonyrefail_site)
# sf::st_write(rhos_tonyrefail_bounding_box, 
#        path(site_folder_path, "bounding_box", "rhos_tonyrefail_bounding_box.kml"))

display_site(rhos_tonyrefail_site, rhos_tonyrefail_bounding_box)

```

### S1 and S2 Granules

### Cleaning the Spatial framework

The Rhos Tonyrefail Phase 1 shapefile was cleaned by removing polygons of <0.01 ha (100m2) and unwanted habitat covers. A total of 5 polygons were removed.   

```{r clean_rhostonyrefail_Phase1_habitat_layer,eval=FALSE}
rhos_raw<-sf::st_read("ST08_Phase1_names.shp")
rhos_clean<-rhos_raw %>% dplyr::filter(!Name == "Mine") %>% dplyr::filter(AREA_HA>=0.01)
sf::st_write(rhos_clean,"Cleaned/ST08_Phase1_names_cleaned2.shp")
```

### Running the zonal stats

```{r run_zonalstats_rhostonyrefail,eval=FALSE}
dirpath<-"ChangeDetection/"
zonal_stats(polys=paste0(dirpath,'Data/Habitat_maps/Wales/Phase_1_corrected_20200814/Cleaned/ST08_Phase1_names_cleaned2.shp')
            polyfield='OBJECTID',
            s2path=paste0(dirpath,'Data/Sentinel2/RhosTonyrefail/'),
            s1path=paste0(dirpath,'Data/Sentinel1/RhosTonyrefail/'),
            outfolder=paste0(dirpath,'Statistics/RhosTonyrefail_Phase1/'),
            sitename='RhosTonyrefail')
```

### Calculate monthly and seasonal summary and change stats
```{r run_changestats_rhostonyrefail,eval=FALSE}
change_stats(sitefile = paste0(dirpath,'Statistics/RhosTonyrefail_Phase1/ZonalStats/RhosTonyrefail_zonal_stats.txt'),
             outfolder = paste0(dirpath,'Statistics/RhosTonyrefail_Phase1/'),
             polygons = paste0(dirpath,'Data/Habitat_maps/Wales/Phase_1_corrected_20200814/Cleaned/ST08_Phase1_names_cleaned2.shp'),
             polyid = 'OBJECTID',
             habclass = 'Name')
```