---
title: "CUU_ChangeDetection Scottish sites"
date: "09/10/2020"
output:
  html_document:
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library(HabChangeDetection)
library(sf)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(tmap)
library(janitor)
library(fs)
library(readxl)
library(magrittr)

#source('habitat_change_detection_functions.R')
```

```{r display_site,include=F,eval=F}
display_site <- function(site_polygons, bounding_box) {
  tmap::tmap_mode("view")
  tmap::tm_shape(bounding_box) +
    tmap::tm_fill(col = "grey", alpha = 0.5) +
    tmap::tm_borders(col = "red", lwd = 2) +
    tmap::tm_shape(site_polygons) +
    tmap::tm_fill(col = "grey", alpha = 0.8) +
    tmap::tm_borders(col = "black", lwd = 1) +
    tmap::tm_basemap(leaflet::providers$OpenStreetMap) +
    tmap::tm_scale_bar()
}
```

```{r display_table, echo=FALSE,include=F}
library(reactable)
display_table <- function(data, rows = nrow(data)) {
  reactable::reactable(data[1:rows, ], fullWidth = FALSE, searchable = TRUE, 
            compact = TRUE, wrap = FALSE, resizable = TRUE,
            defaultColDef = colDef(align = "left"),
            showPageSizeOptions = TRUE, pageSizeOptions = c(10, nrow(data)),
            class = "table")
}
```

```{r get_granule_present_month,include=F}
# at least one S2 granules present per month
get_granule_present_month <- function(granules_dataset){
  granules_dataset %>% 
    dplyr::mutate(year = year(capture_date),
                  month = month(capture_date)) %>% 
    dplyr::group_by(year, month) %>% 
    dplyr::tally() %>% 
    dplyr::full_join(tidyr::crossing(year = seq(2015, 2020), month = seq(1, 12))) %>% 
    dplyr::mutate(n = tidyr::replace_na(n, 0)) %>% 
    dplyr::arrange(year, month) %>% 
    dplyr::mutate(month = month.abb[month]) %>% 
    tidyr::pivot_wider(names_from = month, values_from = n) %>% 
    dplyr::mutate(across(.cols = Jan:Dec, .fns = function(x){if(x > 0){ x / x } else 0})) %>% 
    janitor::adorn_totals(where = "col")
}
```

```{r paths, include=F,eval=F}
site_folder_path <- path("Sites/")
granules_folder_path <- path("ChangeDetection/Data/Sentinel2/")
```

## Copernicus User Uptake - Change detection

Scottish sites:

* HES site 2 (10km^2^) 
* Insh Marshes SSSI, Glenfeshie and HES site 3 (15km^2^): 
* Flanders Moss SSSI and HES site 4 (10km^2^) 
* Skyreburn Grasslands SSSI, Lagganmullan SSSI and HES site 1 (10km^2^) 

Spatial frameworks:

* EUNIS Landcover map 10m raster layer
* 100m^2/1 hectare grid square EUNIS classification


```{r allsites, include=FALSE,eval=F}
scotland_sites <- sf::st_read(path(site_folder_path, "CUU_CD_Scottish_Sites.shp")) %>% 
  janitor::clean_names() %>% 
  dplyr::rename(code = id,
         name = site_name) %>% 
  dplyr::mutate(name = as.character(name)) %>% 
  dplyr::select(name, code)
```


## Create spatial frameworks

```{r spatframework,eval=F}
library(raster)
library(dplyr)
library(APfun)
library(rgdal)

filepath <- 'filepath/'
lookup <- read.csv('EUNIS_habitat_lookup.txt')

sitetifs <- list.files(filepath, pattern='.tif', full.names=T)[c(1,4,7,10)]
if(!dir.exists(paste0(filepath,'Framework2'))){
  dir.create(paste0('Framework2'))
}

#iterate through 4 sites
purrr::map(sitetifs,.f=function(tif){
  gdalUtils::gdal_setInstallation()
  #aggregate to 100m by modal group
  gdalUtils::gdalwarp(srcfile = tif,
                      dstfile = paste0(dirname(tif),'/Framework2/1ha_',basename(tif)), 
                          s_srs = '+init=epsg:27700', t_srs = '+init=epsg:27700', r = 'mode', co = 'COMPRESS=LWZ', overwrite=T,
                      tr = c(100,100))
  # create a blank raster with individual cell ids
  ha_rast <-  raster::raster(paste0(dirname(tif),'/Framework2/1ha_',basename(tif)))
  blank_rast <- ha_rast
  ha_df <- raster::as.data.frame(blank_rast) %>% dplyr::mutate(ID=dplyr::row_number())
  blank_rast$vals <- ha_df$ID
  raster::writeRaster(blank_rast$vals,paste0(dirname(tif),'/Framework2/1ha_blank_',basename(tif)),overwrite=T)
  #create polygon grid shapefile
  APfun::APpolygonize(paste0(dirname(tif),'/Framework2/1ha_blank_',basename(tif)),
                      outFile=paste0(dirname(tif),'/Framework2/1ha_blank_',gsub(basename(tif),pattern='.tif',replacement='.shp')),
                      OSGeoPath  = 'D:/Programs/QGIS',readToMemory = TRUE)
  
  #extract polygon value from raster
  grid_poly <- sf::st_read(paste0(dirname(tif),'/Framework2/1ha_blank_',gsub(basename(tif),pattern='.tif',replacement='.shp')))
  #extract all polygon values
  poly_stat <- raster::extract(ha_rast,grid_poly,progress = "window",df=T)
  
 # Join back into polygons
  names(poly_stat) <-c("ID","class")
  poly_df <- grid_poly %>% 
    dplyr::left_join(poly_stat,by=c('DN'='ID'))  %>%
    dplyr::left_join(lookup[,c('RASTER_CODE','EUNIS_CODE','EUNIS_DESC')],by=c('class'='RASTER_CODE') ) %>% 
    dplyr::select(-class)
  
  sf::st_write(poly_df,paste0(dirname(tif),'/Framework2/1ha_grid_EUNIS',gsub(basename(tif),pattern='.tif',replacement='.shp')),delete_layer = T)

  
})


```

## HES site 2 (10km^2^) 

**S2 Granule:** T30VVK_ORB080, T30VVK_ORB123

* HES site 2 (Peat and moorland; Aerial and ground surveys 2020)

```{r HesSite2_BB, fig.width=8, fig.height=4,eval=F}
hes_site_2 <- scotland_sites %>% dplyr::filter(code == 2)
hes_site_bounding_box <- create_square_bounding_box_polygon(hes_site_2)
# st_write(hes_site_bounding_box, 
#           path(site_folder_path, "bounding_box", "hes_site_2_bounding_box.kml"))
display_site(hes_site_2, hes_site_bounding_box)

```

### S1 and S2 Granules

### Cleaning the Spatial frameworks

```{r hessite2clean,eval=F}

## Spatial framework 1 - 10m raster layer
polys <- sf::st_read('1ha_grid_EUNISELCS_CUU_CD_Scottish_Sites_id_2.shp',quiet=T)

#remove polygons under 100m
polys_sf <- polys %>% dplyr::mutate(area = round(Shape_Area,digits=0)) %>% 
  dplyr::filter(area >100)
print(paste0("Polygons remaining: ", nrow(polys_sf)))

#remove habitats we dont want to monitor
levels(polys$EUNIS_DESC)

habitat_remove <- c("B Roads, local streets and minor roads" ,
                    "Buildings of cities, towns and villages / Low density buildings",
                    "Private roads, publically accessible and restricted"  ,
                    "Road networks" ,
                    "Transport networks and other constructed hard-surfaced areas" )
polys_sf <- polys_sf %>% dplyr::filter(!EUNIS_DESC %in% habitat_remove)
print(paste0("Polygons remaining: ", nrow(polys_sf)))

#write out clean version
if(!dir.exists('Habitat_maps/Scotland/cleaned')}
st_write(polys_sf, 'Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_2_clean.shp', delete_layer=T)

##---------------##

# Spatial framework 2 - 1ha grid

polys2 <- st_read('1ha_grid_EUNISELCS_CUU_CD_Scottish_site_2.shp')

#remove habitats we dont want to monitor or na
unique(polys2$EUNIS_DESC)
habitat_remove <- c("A Road" ,
                    "B Roads, local streets and minor roads",
                    "Buildings of cities, towns and villages / Low density buildings"  ,
                    "Extractive industrial sites / Waste deposits",
                    "Private roads, publically accessible and restricted",
                    "Rail networks" ,
                    "Road networks" ,
                    "Transport networks and other constructed hard-surfaced areas"
)
polys_sf <- polys2 %>% dplyr::filter(!EUNIS_DESC %in% habitat_remove) %>%
  dplyr::filter(!is.na(EUNIS_DESC)) 
print(paste0("Polygons remaining: ", nrow(polys_sf)))

st_write(polys_sf, 'Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_2_clean.shp',delete_layer=T)

```

### Spatial Framework 1 - Running the zonal stats
```{r, eval=F}
dirpath <- 'Scotland/'
## take the polygon shapefile
zonal_stats(polys=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_2_clean.shp"),
            polyfield='Id',
            s2path=paste0(dirpath,"Data/Sentinel2/HES_site_2/"),
            s1path=paste0(dirpath,'Data/Sentinel1/HES_site_2/'),
            outfolder=paste0(dirpath,'Statistics/HES_site_2_EUNIS/'),
            sitename="HES_Site2")

```

### Spatial Framework 1 - Calculate monthly and seasonal summaries and change stats
```{r, eval=F}
change_stats(sitefile=paste0(dirpath,'Statistics/HES_site_2_EUNIS/ZonalStats/HES_Site2_zonal_stats.txt'),
             outfolder=paste0(dirpath,'Statistics/HES_site_2_EUNIS/'),
             polygons=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_2_clean.shp"),
             polyid='Id',habclass='EUNIS_DESC')

```

### Spatial Framework 2 - Running the zonal stats
```{r eval=F}
dirpath <- 'Scotland/'
## take the polygon shapefile
zonal_stats(polys=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_2_clean.shp"),
            polyfield='DN',
            s2path=paste0(dirpath,"Data/Sentinel2/HES_site_2/"),
            s1path=paste0(dirpath,'Data/Sentinel1/HES_site_2/'),
            outfolder=paste0(dirpath,'Statistics/HES_site_2_1haGrid/'),
            sitename="HES_Site2")

```

### Spatial Framework 2 - Calculate monthly and seasonal summaries and change stats
```{r, eval=F}
change_stats(sitefile=paste0(dirpath,'Statistics/HES_site_2_1haGrid/ZonalStats/HES_Site2_zonal_stats.txt'),
             outfolder=paste0(dirpath,'Statistics/HES_site_2_1haGrid/'),
             polygons=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_2_clean.shp"),
             polyid='DN',habclass='EUNIS_DESC')

```

### creating S1 thumbnails - only needed once per site (now incorporated into zonal_stats function)

```{r,eval=F}
s1thumbs(s1path=paste0(dirpath,'Data/Sentinel1/HES_site_2/'),
         out_folder=paste0(dirpath,'Statistics/HES_site_2_EUNIS/ZonalStats/s1/thumbs/'))
   
```

## Insh Marshes SSSI, Glenfeshie and HES site 3 (15km^2^): 

**S2 Granule:** T30VVJ_ORB080, T30VVJ_ORB123

* Insh Marshes SSSI (Floodplain wetlands, managed by pony grazing, renaturalising flood managment)
* Glenfeshie (Cairngorms SSSI) (woodland recovering from overgrazing)
* HES site 3

```{r inshsite,  fig.width=8, fig.height=4,eval=F}
insh_glenfeshie_hes3_site <- scotland_sites %>% filter(code == 1)

insh_glenfeshie_hes3_bounding_box <- create_square_bounding_box_polygon(insh_glenfeshie_hes3_site)
# sf::st_write(insh_glenfeshie_hes3_bounding_box, 
#           path(site_folder_path, "bounding_box", "insh_glenfeshie_hes3_bounding_box.kml"))

display_site(insh_glenfeshie_hes3_site, insh_glenfeshie_hes3_bounding_box)

```

### S1 and S2 Granules

### Cleaning the Spatial framework

```{r inshclean,eval=F}

polys <- sf::st_read('1ha_grid_EUNISELCS_CUU_CD_Scottish_Sites_id_1.shp',quiet=T)

#remove polygons under 100m
polys_sf <- polys %>% dplyr::mutate(area = round(Shape_Area,digits=0)) %>% 
  dplyr::filter(area >100)
print(paste0("Polygons remaining: ", nrow(polys_sf)))

#remove habitats we dont want to monitor
levels(polys$EUNIS_DESC)

habitat_remove <- c("A Road" ,
                    "B Roads, local streets and minor roads",
                    "Buildings of cities, towns and villages / Low density buildings"  ,
                    "Extractive industrial sites / Waste deposits",
                    "Private roads, publically accessible and restricted",
                    "Rail networks" ,
                     "Road networks" ,
                    "Transport networks and other constructed hard-surfaced areas"
                    )
polys_sf <- polys_sf %>% dplyr::filter(!EUNIS_DESC %in% habitat_remove)
print(paste0("Polygons remaining: ", nrow(polys_sf)))

#write out clean version
sf::st_write(polys_sf, 'ELCS_Scottish_site_1_clean.shp',delete_layer=T)

##---------------##

# Spatial framework 2 - 1ha grid

polys2 <- sf::st_read('1ha_grid_EUNISELCS_CUU_CD_Scottish_Sites_id_1.shp')

#remove habitats we dont want to monitor or na
unique(polys2$EUNIS_DESC)

habitat_remove <- c("A Road" ,
                    "B Roads, local streets and minor roads",
                    "Buildings of cities, towns and villages / Low density buildings"  ,
                    "Extractive industrial sites / Waste deposits",
                    "Private roads, publically accessible and restricted",
                    "Rail networks" ,
                     "Road networks" ,
                    "Transport networks and other constructed hard-surfaced areas"
                    )
polys_sf <- polys2 %>% dplyr::filter(!EUNIS_DESC %in% habitat_remove) %>%
  dplyr::filter(!is.na(EUNIS_DESC)) 
print(paste0("Polygons remaining: ", nrow(polys_sf)))

sf::st_write(polys_sf, '1haGrid_ELCS_Scottish_site_1_clean.shp',delete_layer=T)

```

### Spatial Framework 1 - Running the zonal stats

```{r, eval=F}
dirpath <- 'Scotland/'
## take the polygon shapefile
zonal_stats(polys=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_1_clean.shp"),
            polyfield='Id',
            s2path=paste0(dirpath,"Data/Sentinel2/InshMarshes/"),
            s1path=paste0(dirpath,'Data/Sentinel1/InshMarshes/'),
            outfolder=paste0(dirpath,'Statistics/InshMarshes_EUNIS/'),
            sitename="InshMarshes")
```


### Spatial Framework 1 -  Calculate monthly and seasonal summaries and change stats
```{r, eval=F}
change_stats(sitefile=paste0(dirpath,'Statistics/InshMarshes_EUNIS/ZonalStats/InshMarshes_zonal_stats.txt'),
             outfolder=paste0(dirpath,'Statistics/InshMarshes_EUNIS/'),
             polygons=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_1_clean.shp"),
             polyid='Id',habclass='EUNIS_DESC')

```


### Spatial Framework 2 - Running the zonal stats

```{r, eval=F}
## take the polygon shapefile
zonal_stats(polys=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_1_clean.shp"),
            polyfield='DN',
            s2path=paste0(dirpath,"Data/Sentinel2/InshMarshes/"),
            s1path=paste0(dirpath,'Data/Sentinel1/InshMarshes/'),
            outfolder=paste0(dirpath,'Statistics/InshMarshes_1haGrid/'),
            sitename="InshMarshes")
```


### Spatial Framework 2 -  Calculate monthly and seasonal summaries and change stats
```{r, eval=F}
change_stats(sitefile=paste0(dirpath,'Statistics/InshMarshes_1haGrid/ZonalStats/InshMarshes_zonal_stats.txt'),
             outfolder=paste0(dirpath,'Statistics/InshMarshes_1haGrid/'),
             polygons=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_1_clean.shp"),
             polyid='DN',habclass='EUNIS_DESC')

```

### creating S1 thumbnails - only needed once per site  (now incorporated into zonal_stats function)

```{r,eval=F}
s1thumbs(s1path=paste0(dirpath,'Data/Sentinel1/InshMarshes/'),
         out_folder=paste0(dirpath,'Statistics/InshMarshes_EUNIS/ZonalStats/s1/thumbs/'))

   
```

## Flanders Moss SSSI and HES site 4 (10km^2^) 

**S2 Granule:** T30VVH_ORB080, T30VVH_ORB123

* Flanders Moss SSSI (Raised bog and bog woodland features, in unfavourable but recovering condition, managed by controlling scrub and maintaining water table)
* HES site 4 (Arable, recreation and woodland)

```{r flanderssite,fig.width=8, fig.height=4,eval=F}
flanders_moss_hes4_site <- scotland_sites %>% filter(code == 3) 

flanders_moss_hes4_bounding_box <- create_square_bounding_box_polygon(flanders_moss_hes4_site)
# sf::st_write(flanders_moss_hes4_bounding_box, 
#         path(site_folder_path, "bounding_box", "flanders_moss_hes4_bounding_box.kml"))
display_site(flanders_moss_hes4_site, flanders_moss_hes4_bounding_box )

```

### S1 and S2 Granules

### Cleaning the Spatial framework

```{r flandersclean,eval=F}

polys <- sf::st_read('ELCS_Scottish_site_3.shp',quiet=T)

#remove polygons under 100m
polys_sf <- polys %>% dplyr::mutate(area = round(Shape_Area,digits=0)) %>% 
  dplyr::filter(area >100)
print(paste0("Polygons remaining: ", nrow(polys_sf)))

#remove habitats we dont want to monitor
levels(polys$EUNIS_DESC)

habitat_remove <- c("A Road" ,
                    "B Roads, local streets and minor roads",
                    "Buildings of cities, towns and villages / Low density buildings"  ,
                    "Private roads, publically accessible and restricted",
                    "Road networks" ,
                    "Transport networks and other constructed hard-surfaced areas"
                    )
polys_sf <- polys_sf %>% dplyr::filter(!EUNIS_DESC %in% habitat_remove)
print(paste0("Polygons remaining: ", nrow(polys_sf)))

#write out clean version
sf::st_write(polys_sf, 'Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_3_clean.shp',delete_layer=T)
##---------------##

# Spatial framework 2 - 1ha grid

polys2 <- sf::st_read('1ha_grid_EUNISELCS_Scottish_site_3.shp')
#remove habitats we dont want to monitor or na
unique(polys2$EUNIS_DESC)

habitat_remove <- c("A Road" ,
                    "B Roads, local streets and minor roads",
                    "Buildings of cities, towns and villages / Low density buildings"  ,
                    "Private roads, publically accessible and restricted",
                    "Road networks" ,
                    "Transport networks and other constructed hard-surfaced areas"
                    )
polys_sf <- polys2 %>% dplyr::filter(!EUNIS_DESC %in% habitat_remove) %>%
  dplyr::filter(!is.na(EUNIS_DESC)) 
print(paste0("Polygons remaining: ", nrow(polys_sf)))

sf::st_write(polys_sf, 'Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_3_clean.shp',delete_layer=T)

```

### Spatial Framework 1 - Running the zonal stats

```{r, eval=F}
dirpath <- 'Scotland/'
## take the polygon shapefile
zonal_stats(polys=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_3_clean.shp"),
            polyfield='Id',
            s2path=paste0(dirpath,"Data/Sentinel2/FlandersMoss/"),
            s1path=paste0(dirpath,'Data/Sentinel1/FlandersMoss/'),
            outfolder=paste0(dirpath,'Statistics/FlandersMoss_EUNIS/'),
            sitename="FlandersMoss")
```


### Spatial Framework 1 -  Calculate monthly and seasonal summaries and change stats
```{r, eval=F}
change_stats(sitefile=paste0(dirpath,'Statistics/FlandersMoss_EUNIS/ZonalStats/FlandersMoss_zonal_stats.txt'),
             outfolder=paste0(dirpath,'Statistics/FlandersMoss_EUNIS/'),
             polygons=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_3_clean.shp"),
             polyid='Id',habclass='EUNIS_DESC')

```


### Spatial Framework 2 - Running the zonal stats

```{r, eval=F}
## take the polygon shapefile
zonal_stats(polys=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_3_clean.shp"),
            polyfield='DN',
            s2path=paste0(dirpath,"Data/Sentinel2/FlandersMoss/"),
            s1path=paste0(dirpath,'Data/Sentinel1/FlandersMoss/'),
            outfolder=paste0(dirpath,'Statistics/FlandersMoss_1haGrid/'),
            sitename="FlandersMoss")
```


### Spatial Framework 2 -  Calculate monthly and seasonal summaries and change stats
```{r, eval=F}
change_stats(sitefile=paste0(dirpath,'Statistics/FlandersMoss_1haGrid/ZonalStats/FlandersMoss_zonal_stats.txt'),
             outfolder=paste0(dirpath,'Statistics/FlandersMoss_1haGrid/'),
             polygons=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_3_clean.shp"),
             polyid='DN',habclass='EUNIS_DESC')

```


### creating S1 thumbnails - only needed once per site  (now incorporated into zonal_stats function)

```{r,eval=F}
s1thumbs(s1path=paste0(dirpath,'Data/Sentinel1/FlandersMoss/'),
         out_folder=paste0(dirpath,'Statistics/FlandersMoss_EUNIS/ZonalStats/s1/thumbs/'))

   
```

## Skyreburn Grasslands SSSI, Lagganmullan SSSI and HES site 1 (10km^2^) 

**S2 Granule:** T30UVF_ORB080, T30UVF_ORB123

* Skyreburn Grasslands, SSSI (lowland neutral grassland and fen meadow, grazing regime)
* Lagganmullan, SSSI (fen meadow without active management; NVC survey 2019.)
* HES site 1 

```{r skyre,  fig.width=8, fig.height=4,eval=F}
skyre_laggan_hes1_site <- scotland_sites %>% filter(code == 4)
skyre_laggan_hes1_bounding_box <- create_square_bounding_box_polygon(skyre_laggan_hes1_site)
# sf::st_write(skyre_laggan_hes1_bounding_box, 
#         path(site_folder_path, "bounding_box", "skyre_laggan_hes1_bounding_box.kml"))
display_site(skyre_laggan_hes1_site, skyre_laggan_hes1_bounding_box)

```


### S1 and S2 Granules

### Cleaning the Spatial framework

```{r skyreclean,eval=F}

polys <- sf::st_read('ELCS_Scottish_site_4.shp',quiet=T)

#remove polygons under 100m
polys_sf <- polys %>% dplyr::mutate(area = round(Shape_Area,digits=0)) %>% 
  dplyr::filter(area >100)
print(paste0("Polygons remaining: ", nrow(polys_sf)))

#remove habitats we dont want to monitor
levels(polys$EUNIS_DESC)

habitat_remove <- c("A Road" ,
                    "B Roads, local streets and minor roads",
                    "Buildings of cities, towns and villages / Low density buildings"  ,
                    "Coastal dunes and sandy shores",
                    "Coastal saltmarshes and saline reedbeds" ,
                     "Coastal shingle" ,
                    "Littoral rock and other hard substrata" ,
                    "Littoral sediments"  ,
                    "Marine habitats" ,
                    "Private roads, publically accessible and restricted",
                    "Road networks" ,
                     "Transport networks and other constructed hard-surfaced areas" 
                    )
polys_sf <- polys_sf %>% dplyr::filter(!EUNIS_DESC %in% habitat_remove)
print(paste0("Polygons remaining: ", nrow(polys_sf)))

#write out clean version
sf::st_write(polys_sf, 'Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_4_clean.shp',delete_layer=T)
##---------------##

# Spatial framework 2 - 1ha grid

polys2 <- sf::st_read('Habitat_maps/Scotland/Framework2/1ha_grid_EUNISELCS_Scottish_site_4.shp')
#remove habitats we dont want to monitor or na
unique(polys2$EUNIS_DESC)

habitat_remove <- c("A Road" ,
                    "B Roads, local streets and minor roads",
                    "Buildings of cities, towns and villages / Low density buildings"  ,
                    "Private roads, publically accessible and restricted",
                    "Road networks" ,
                    "Transport networks and other constructed hard-surfaced areas"
)
polys_sf <- polys2 %>% dplyr::filter(!EUNIS_DESC %in% habitat_remove) %>%
  dplyr::filter(!is.na(EUNIS_DESC))
print(paste0("Polygons remaining: ", nrow(polys_sf)))

sf::st_write(polys_sf, 'Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_4_clean.shp',delete_layer=T)

```

### Spatial Framework 1 - Running the zonal stats

```{r, eval=F}
## take the polygon shapefile
zonal_stats(polys=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_4_clean.shp"),
            polyfield='Id',
            s2path=paste0(dirpath,"Data/Sentinel2/Skyreburn/"),
            s1path=paste0(dirpath,'Data/Sentinel1/Skyreburn/'),
            outfolder=paste0(dirpath,'Statistics/Skyreburn_EUNIS/'),
            sitename="Skyreburn")
```


### Spatial Framework 1 -  Calculate monthly and seasonal summaries and change stats
```{r, eval=F}
change_stats(sitefile=paste0(dirpath,'Statistics/Skyreburn_EUNIS/ZonalStats/Skyreburn_zonal_stats.txt'),
             outfolder=paste0(dirpath,'Statistics/Skyreburn_EUNIS/'),
             polygons=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/ELCS_Scottish_site_4_clean.shp"),
             polyid='Id',habclass='EUNIS_DESC')

```


### Spatial Framework 2 - Running the zonal stats

```{r, eval=F}
## take the polygon shapefile
zonal_stats(polys=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_4_clean.shp"),
            polyfield='DN',
            s2path=paste0(dirpath,"Data/Sentinel2/Skyreburn/"),
            s1path=paste0(dirpath,'Data/Sentinel1/Skyreburn/'),
            outfolder=paste0(dirpath,'Statistics/Skyreburn_1haGrid/'),
            sitename="Skyreburn")
```


### Spatial Framework 2 -  Calculate monthly and seasonal summaries and change stats
```{r, eval=F}
change_stats(sitefile=paste0(dirpath,'Statistics/Skyreburn_1haGrid/ZonalStats/Skyreburn_zonal_stats.txt'),
             outfolder=paste0(dirpath,'Statistics/Skyreburn_1haGrid/'),
             polygons=paste0(dirpath,"Data/Habitat_maps/Scotland/cleaned/1haGrid_ELCS_Scottish_site_4_clean.shp"),
             polyid='DN',habclass='EUNIS_DESC')

```


### creating S1 thumbnails - only needed once per site  (now incorporated into zonal_stats function)

```{r,eval=F}
s1thumbs(s1path=paste0(dirpath,'Data/Sentinel1/Skyreburn/'),
         out_folder=paste0(dirpath,'Statistics/Skyreburn_EUNIS/ZonalStats/s1/thumbs/'))

   
```
