---
title: "CUU_ChangeDetection Preparing data for the app"
date: '09/10/2020'
author: 'JNCC'
licence: 'MIT licence'
output:
  html_document:
    df_print: paged
    css: style.css
    includes:
      before_body: header.html
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Creating thumbnail legends

The 'legend_thumbnail' code in the 'habitat_change_detection_functions.R' script will create images containing the legends for the indices thumbnails created in the zonal stats function. This is run using the below code. The legends for the RVI and RVIv are slightly different with edges tacked on in order to show that the min and max values are beyond the break range of the colour palette stretch.This was in order to display the variation in the data better.

```{r legendprep,eval=F}

lookup <- tibble::tribble(~index,~breaks,~palette,~contlabel,~fixedlabel,
                          "NDVI", seq(-1,1,by=0.2),terrain.colors(10,rev=T),
                          c("-1","","-0.6","","","0","","","0.6","","1"),
                          c("-1","","","","","0","","","","1"),
                          "NDWI",seq(-1,1,by=0.2),"Blues",
                          c("-1","","-0.6","","","0","","","0.6","","1"),
                          c("-1","","","","","0","","","","1"),
                          "NDMI",seq(-1,1,by=0.2),"Purples",
                          c("-1","","-0.6","","","0","","","0.6","","1"),
                          c("-1","","","","","0","","","","1"),
                          "RVI", seq(0.9,1.7,by=0.1),"viridis",
                          c("","1","","","1.3","","","1.6",""),
                          c("","1","","","1.3","","1.6",""),
                          "RVIv",seq(-0.2,4.2,by=0.2),c("#440154FF",viridis(10),"#FDE725FF"),
                          c("","0","","","","","1","","","","","2","","","","","3","","","","","4",""), 
                          c("0","","","","","1","","","","","2","","","","","3","","","","","4","","","","5"))

# load in raster template layers for the relevant indices
#s2
r <- "SEN2_20150628_lat54lon368_T30UVE_ORB077_utm30n_osgb_vmsk_sharp_rad_srefdem_stdsref_CorsBodeilio_msk_NDVI.tif"
#s1
r <- "S1B_20161002_30_asc_175722_175747_VVVH_G0_GB_OSGB_RTCK_SpkRL_CorsBodeilio_RVI.tif"

#iterate through the indices
purrr::map(lookup$index, .f=function(ind){
  indexbreak <- lookup %>% dplyr::filter(index==ind)

legend_thumbnail(r, index=indexbreak$index,breaks=unlist(indexbreak$breaks),palette=unlist(indexbreak$palette),
                 outfolder='./Index_legends/',
                 labels=unlist(indexbreak$contlabel),continuous=T)
})


```


## Make edits to the text files

This runs through two file edit functions:

* fileEdit - Function to edit the zonal stats txt files to remove additional rownames column and add a monthdate column
* fileEdit2 - Function to write the habitat names into the zonal stats txt files


```{r filedits,eval=F}
source('./HabChangeDetection/R/')
## Appy this fileEdit function across all files - edit the zonal stats txt files to Remove the additional rownames column and add in the monthdate column if it's missing
fls <- list.files("./data/", full.names = T, recursive = T)
## Subset the files if necessary
# fls <- fls[grepl(fls, pattern = "peak")]
out <- pbapply::pblapply(fls, fileEdit)


## Appy this fileEdit2 function across all the zonal stats files - writes the habitat names into the zonal stats txt files
fls <- list.files("./data", full.names = T, recursive = T)
## Subset if necessary
#fls <- fls[grepl(fls, pattern = "peak")]
config <- yaml::read_yaml('./config.yaml')
out <- pbapply::pblapply(fls, fileEdit2, config)

```


## Flag Change in the summarised statistics

Use the polygon and habitat level statistics to calculate change

```{r flagchange,eval=F}

## List zonal stats folders
files <- c('./data/corsbodeilio/phase1/monthly/zonal_stats/',
           './data/corsbodeilio/phase1/seasonal/zonal_stats/',
           './data/rhostonyrefail/phase1/monthly/zonal_stats/',
           './data/rhostonyrefail/phase1/seasonal/zonal_stats/',
           './data/darkpeak/phi/monthly/zonal_stats/',
           './data/darkpeak/phi/seasonal/zonal_stats/',
           './data/malvernhills/phi/monthly/zonal_stats/',
           './data/malvernhills/phi/seasonal/zonal_stats/',
           './data/flandersmoss/eunis/monthly/zonal_stats/',
           './data/flandersmoss/eunis/seasonal/zonal_stats/',
           './data/flandersmoss/hagrid/monthly/zonal_stats/',
           './data/flandersmoss/hagrid/seasonal/zonal_stats/',
           './data/inshmarshes/eunis/monthly/zonal_stats/',
           './data/inshmarshes/eunis/seasonal/zonal_stats/',
           './data/inshmarshes/hagrid/monthly/zonal_stats/',
           './data/inshmarshes/hagrid/seasonal/zonal_stats/',
           './data/skyreburn/eunis/monthly/zonal_stats/',
           './data/skyreburn/eunis/seasonal/zonal_stats/',
           './data/skyreburn/hagrid/monthly/zonal_stats/',
           './data/skyreburn/hagrid/seasonal/zonal_stats/',
           './data/hes2/eunis/monthly/zonal_stats/',
           './data/hes2/eunis/seasonal/zonal_stats/',
           './data/hes2/hagrid/monthly/zonal_stats/',
           './data/hes2/hagrid/seasonal/zonal_stats/')

## List indices
indices <- c('NDVI', 'NDWI', 'NDMI', 'RVI', 'RVIv')

# run flag change function
flag_change(files, indices)

```

## Create an index of names

Create a list of “raw” S1/S2 tifs, and index tifs, to use to query AWS

```{r indexlist,eval=F}
## Load packages
require(tidyverse)

## List all files in the indices folder
flsList <- list.files("../indices", full.names = T, recursive = T)

## Remove the leading filepath
flsList <- gsub(flsList, pattern = "\\Q../\\E", replacement = "")

## Create a tibble of paths
flsList <- tibble::tibble(PATH = flsList)

## Extract the site, index, and file names to add to separate columns
flsList <- flsList %>%
  dplyr::mutate(SITE = stringr::str_split(PATH, pattern = "/", simplify = T)[,2],
         INDEX = stringr::str_split(PATH, pattern = "/", simplify = T)[,4],
         FILE = stringr::str_split(PATH, pattern = "/", simplify = T)[,5])

## Write the list of imagery out
readr::write_csv(flsList, "./imageryList.csv")


```

